---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(dplyr)
library(ggplot2)

data<-read.csv("activity.csv")


```





## What is mean total number of steps taken per day?
```{r}
j<-data %>% group_by(date) %>% summarise(no_steps=sum(steps,na.rm = TRUE))
hist(j$no_steps,col="blue",xlab="No of Steps",ylab = "Average no of Steps",main="Total no of steps per day ",ylim = c(0,35))

print(paste("Median of total number of steps per day",median(j$no_steps)))
print(paste("Mean of total number of steps per day",mean(j$no_steps)))

```



## What is the average daily activity pattern?

```{r}
j1<-data %>% group_by(interval) %>% summarise(no_steps=mean(steps,na.rm = TRUE))

plot(j1$interval,j1$no_steps,type = 'l',col="blue",xlab="Interval",ylab = "Average no of Steps",main="Average no of steps over intervals")

m<-j1$interval[which.max(j1$no_steps)]

print(paste("The 5 minute interval which contains the maximum number of steps is",m))


```




## Imputing missing values

#### Replacing the missing values with average no of steps with respective to that interval which has been calculated in the previous step(j1)

```{r}

final<-data
for( i in  1:nrow(final))
{
  if (is.na(final[i,1])==TRUE)
  {
    final[i,1]<-j1[which(final[i,3]==j1[,1]),2]
  }
}

j2<-final %>% group_by(date) %>% summarise(no_steps=sum(steps,na.rm = TRUE))

hist(j$no_steps,col="blue",xlab="No of Steps",ylab = "Average no of Steps",main="Total no of steps per day ",ylim = c(0,35))

namean<-data.frame(mean=c(mean(j$no_steps),mean(j2$no_steps)),median=c(median(j$no_steps),median(j2$no_steps)))

row.names(namean)<-c("With Missing Values","Without Missing values")

namean


```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
final[,2]<-as.Date(final[,2])

final<-mutate(final,days= weekdays(final$date))

for( i in 1:nrow(final))
{
  if ( final[i,4] %in% c("Saturday","Sunday"))
  {
    final[i,4]<-"Weekend"
  }else
  {
    final[i,4]<-"Weekday"
  }
}

j3<-final %>% group_by(interval,days) %>% summarise(no_steps=mean(steps,na.rm = TRUE))


g<-ggplot(j3,aes(interval,no_steps))
g+geom_line()+facet_grid(days~.)+ggtitle("Average Number of Steps - Weekday vs. Weekend") 


```


